### 2016-05-24 paper/icfp16-cameraready/related.tex
misc

Lerner \etal~\cite{lerner_searching_2007} attempt to localize the type
error and suggest a fix by replacing expressions (or removing them
entirely) with alternatives based on the surrounding program context.
%
Chen and Erwig~\cite{chen_counter-factual_2014} use a variational type
system to allow for the possibility of changing an expression's type,
and search for an expression whose type can be changed such that type
inference would succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner, who searches for changes at the
value-level, by searching at the type level Chen's search is complete due
the finite universe of types in a program.
%
Neubauer and Thiemann~\cite{neubauer_discriminative_2003} present a
decidable type system based on discriminative sum types, in which all
terms are typeable and type derivations contain all type errors in a
program. They then use the typing derivation to slice out the parts of
the expression related to each error.
%
Zhang and Myers~\cite{zhang_toward_2014} present an algorithm for
identifying the most likely culprit in a system of unsatisfiable
constraints (\eg type equalities), based on Bayesian reasoning.
%
Pavlinovic \etal~\cite{pavlinovic_finding_2014} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}

@@

\paragraph{Running Ill-Typed Programs}
\label{sec:running-ill-typed}
Vytiniotis \etal~\cite{vytiniotis_equality_2012} extend the Haskell
compiler GHC to support compiling ill-typed programs, but their intent
is rather different from ours. Their goal was to allow programmers to
incrementally test refactorings, which often cause type errors in
distant functions. They replace any expression that fails to type
check with a \emph{runtime} error, but do not check types
at runtime.

\paragraph{Testing}
\label{sec:testing}
\nanomaly is at its heart a test generator, and builds on a rich line of
work.
%
Our use of holes to represent unknown values is inspired by the work of
Runciman, Naylor, and Lindblad~\cite{runciman_smallcheck_2008,naylor_finding_2007,lindblad_property_2007},
%
who use lazy evaluation to drastically reduce the search space for
exhaustive test generation, by grouping together equivalent inputs by
the set of values they force. An exhaustive search is complete (up to
the depth bound), if a witness exists it will be found, but due to the
exponential blowup in the search space the depth bound can be quite
limited without advanced grouping and filtering techniques.
%
Our search is not exhaustive; instead we use random generation to fill
in holes on demand.
%
Random test generation~\cite{claessen_quickcheck:_2000,csallner_jcrasher:_2004,pacheco_feedback-directed_2007}
%
is by its nature incomplete, but is able to check larger inputs than
exhaustive testing as a result.

Instead of enumerating values, which may trigger the same path through
the program, one might enumerate paths. 
%
Dynamic-symbolic execution~\cite{godefroid_dart:_2005,cadar_klee:_2008,tillmann_pex_2008}
%
combines symbolic execution (to track which path a given input
triggers) with concrete execution (to ensure failures are
not spurious). The system collects a path condition during execution,
which tracks symbolically what conditions must be met to trigger the
current path. Upon successfully completing a test run, it negates the
path condition and queries a solver for another set of inputs that
satisfy the negated path condition, \ie inputs that will not trigger the
same path. Thus, it can prune the search space
much faster than techniques based on enumerating values, but is limited
by the expressiveness of the underlying solver. Our operational
semantics is amenable to dynamic-symbolic execution, one would just need
to collect the path condition and replace our implementation of \gensym
by a call to the solver. We chose to use lazy, random generation instead
because it is efficient, and % does not incur the overhead of an external solver, and
produces high coverage for our domain of novice programs.

% \begin{itemize}
% \item \cite{claessen_quickcheck:_2000}
% \item 
% \item \cite{godefroid_dart:_2005}
% \item \cite{cadar_klee:_2008}
% \item \cite{tillmann_pex_2008}
% \end{itemize}

\paragraph{Program Exploration}
Perera \etal~\cite{perera_functional_2012} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} -- containing holes -- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration. 
Perera's system focuses on backward exploration; in
contrast, our visualization supports forward \emph{and} backward
exploration, though our backward steps are more limited than
Perera's. Specifically, we do not support selecting a value and
inserting the intermediate terms that preceded it while ignoring
unrelated computation steps; this would be interesting future work.

>>>>

Lerner \etal~\cite{Lerner2007-dt} attempt to localize the type
error and suggest a fix by replacing expressions (or removing them
entirely) with alternatives based on the surrounding program context.
%
Chen and Erwig~\cite{Chen2014-gd} use a variational type
system to allow for the possibility of changing an expression's type,
and search for an expression whose type can be changed such that type
inference would succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner, who searches for changes at the
value-level, by searching at the type level Chen's search is complete due
the finite universe of types in a program.
%
Neubauer and Thiemann~\cite{Neubauer2003-xv} present a
decidable type system based on discriminative sum types, in which all
terms are typeable and type derivations contain all type errors in a
program. They then use the typing derivation to slice out the parts of
the expression related to each error.
%
Zhang and Myers~\cite{Zhang2014-lv} present an algorithm for
identifying the most likely culprit in a system of unsatisfiable
constraints (\eg type equalities), based on Bayesian reasoning.
%
Pavlinovic \etal~\cite{Pavlinovic2014-mr} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}

@@

\paragraph{Running Ill-Typed Programs}
\label{sec:running-ill-typed}
Vytiniotis \etal~\cite{Vytiniotis2012-gh} extend the Haskell
compiler GHC to support compiling ill-typed programs, but their intent
is rather different from ours. Their goal was to allow programmers to
incrementally test refactorings, which often cause type errors in
distant functions. They replace any expression that fails to type
check with a \emph{runtime} error, but do not check types
at runtime.

\paragraph{Testing}
\label{sec:testing}
\nanomaly is at its heart a test generator, and builds on a rich line of
work.
%
Our use of holes to represent unknown values is inspired by the work of
Runciman, Naylor, and Lindblad~\cite{Runciman2008-ka,Naylor2007-mi,Lindblad2007-oy},
%
who use lazy evaluation to drastically reduce the search space for
exhaustive test generation, by grouping together equivalent inputs by
the set of values they force. An exhaustive search is complete (up to
the depth bound), if a witness exists it will be found, but due to the
exponential blowup in the search space the depth bound can be quite
limited without advanced grouping and filtering techniques.
%
Our search is not exhaustive; instead we use random generation to fill
in holes on demand.
%
Random test generation~\cite{Claessen2000-lj,Csallner2004-bf,Pacheco2007-at}
%
is by its nature incomplete, but is able to check larger inputs than
exhaustive testing as a result.

Instead of enumerating values, which may trigger the same path through
the program, one might enumerate paths. 
%
Dynamic-symbolic execution~\cite{Godefroid2005-am,Cadar2008-kg,Tillmann2008-qc}
%
combines symbolic execution (to track which path a given input
triggers) with concrete execution (to ensure failures are
not spurious). The system collects a path condition during execution,
which tracks symbolically what conditions must be met to trigger the
current path. Upon successfully completing a test run, it negates the
path condition and queries a solver for another set of inputs that
satisfy the negated path condition, \ie inputs that will not trigger the
same path. Thus, it can prune the search space
much faster than techniques based on enumerating values, but is limited
by the expressiveness of the underlying solver. Our operational
semantics is amenable to dynamic-symbolic execution, one would just need
to collect the path condition and replace our implementation of \gensym
by a call to the solver. We chose to use lazy, random generation instead
because it is efficient, and % does not incur the overhead of an external solver, and
produces high coverage for our domain of novice programs.

% \begin{itemize}
% \item \cite{Claessen2000-lj}
% \item 
% \item \cite{Godefroid2005-am}
% \item \cite{Cadar2008-kg}
% \item \cite{Tillmann2008-qc}
% \end{itemize}

\paragraph{Program Exploration}
Perera \etal~\cite{Perera2012-dy} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} -- containing holes -- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration. 
Perera's system focuses on backward exploration; in
contrast, our visualization supports forward \emph{and} backward
exploration, though our backward steps are more limited than
Perera's. Specifically, we do not support selecting a value and
inserting the intermediate terms that preceded it while ignoring
unrelated computation steps; this would be interesting future work.

### 2016-05-24 paper/icfp16-cameraready/semantics.tex
init paper/icfp16-cameraready
misc

Next, we formalize the notion of type error witnesses as follows.
%
First, we define a core calculus within which we will work~(\S~\ref{sec:syntax}).
%
Second, we develop a (non-deterministic) operational semantics
for ill-typed programs that precisely defines the notion
of a \emph{witness}~(\S~\ref{sec:semantics}).
%
Third, we formalize and prove a notion of \emph{generality} for
witnesses, which states, intuitively, that if we find a
single witness then for \emph{every possible} type
assignment there exist inputs that are guaranteed to make
the program ``go wrong''~(\S~\ref{sec:soundness}).
%
Finally, we refine the operational semantics into a
\emph{search procedure} that returns concrete (general)
witnesses for ill-typed programs~\S~(\ref{sec:search-algorithm}).
%
We have formalized and tested our semantics and generality theorem
in \textsc{PLT-Redex}~\cite{felleisen2009semantics}.

@@

\paragraph{Holes}
\label{sec:holes}
%
Recall that a key challenge in our setting is to find witnesses
that are meaningful and do not arise from choosing values from
irrelevant types.
%
We solve this problem be equipping our term language with a
notion of a \emph{hole}, written \vhole{\thole}, which represents
an \emph{unconstrained} value $\ehole$ that may be replaced with
\emph{any} value of an unknown type \thole.
%
Intuitively, the type holes \thole\ can be viewed as type variables
that we will \emph{not} generalize over.
%
A \emph{normalized} value is one that is not a hole,
but which may internally contain holes.
%
For example
$\vnode{\thole}{\vhole{\thole}}{\vleaf{\thole}}{\vleaf{\thole}}$ is a
normalized value.

\paragraph{Substitutions}
%
Our semantics ensure the generality of witnesses by incrementally
\emph{refining} holes, filling in just as much information as is
needed locally to make progress (inspired by the manner in
which SmallCheck uses lazy evaluation~\cite{runciman_smallcheck_2008}).
%
We track how the holes are incrementally filled in, by using
value (resp.\ type) \emph{substitutions} $\vsu$ (resp. $\tsu$)
that map value (resp.\ type) holes to values (resp.\ types).
%
The substitutions let us ensure that we consistently instantiate each
hole with the same (partially defined) value or type, regardless of the
multiple contexts in which the hole appears.
%
This ensures we can report a concrete (and general) witness for any
(dynamically) discovered type errors.

@@

\begin{figure}[ht]
$$
\begin{array}{lcll}
\gensym       & :   & t \times \tsu \rightarrow v \\
\gen{\thole}{\tsu}  & \defeq  & \gen{\subst{\tsu}{\thole}}{\tsu}, &  \text{if } \thole \in dom(\tsu) \\
\gen{\tint}{\tsu}   & \defeq  & n, &  \text{non-deterministic} \\
\gen{\tbool}{\tsu}  & \defeq  & b, &  \text{non-deterministic} \\
\gen{\tprod{t_1}{t_2}}{\tsu}  & \defeq  & \epair{\gen{t_1}{\tsu}}{\gen{t_2}{\tsu}} & \\ % \begin{subarray}{l} v_1 = \gen{t_1}{\tsu}\\ v_2 = \gen{t_2}{\tsu}\end{subarray} \\
\gen{\ttree{t}}{\tsu}  & \defeq  & tr, &  \text{non-deterministic} \\
\gen{\tfun}{\tsu}   & \defeq & \efun{x}{\vhole{\thole}}, &  \text{\ehole and \thole are fresh} \\
\gen{\thole}{\tsu}  & \defeq & \vhole{\thole}, & \text{\ehole is fresh} \\
\end{array}
$$
\caption{Generating values}
\label{fig:gen}
\end{figure}

@@

\paragraph{Recursion}
%
% \RJ{this para appears out of nowhere --- non-sequitur}
Our semantics lacks a built-in $\mathtt{fix}$ construct
for defining recursive function, which may surprise
the reader.
%
Fixed-point operators often cannot be typed in static type
systems, but our system would simply approximate its type
as $\tfun$, apply it, and move along with evaluation.
%
Thus we do not need a built-in recursion construct, we
can use any of the standard fixed-point operators.

@@

\paragraph{Preservation}
%
We prove two preservation lemmas. First, we show that each evaluation
step refines the partial input type of $f$, thus preserving type
compatibility.
%
\begin{lem}
\label{lem:vsu-ext}
If  $\trace \defeq \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},\ldots$
and $\trace'$ is a single-step extension of $\trace$
and $\ptype{\trace}{f} \neq \ptype{\trace'}{f}$
% and $\resolve{\thole}{\tsu} \neq \resolve{\thole}{\tsu'}$
then $\tsu' = \tsu[\thole_1 \mapsto t_1] \ldots [\thole_n \mapsto t_n]$.
\end{lem}
\begin{proof}
  By case analysis on the evaluation rules.
  %
  $\thole$ does not change, so if the partial input types differ then
  $\tsu \neq \tsu'$.
  %
  Note that only \forcesym can change \tsu, and it can only do so
  via \unifysym, which can only extend \tsu.
  % and only the first case,
  % when we extend \vsu\ with $[\ehole \mapsto v]$,
  % does.
  %
  % \ES{fill in cases in the appendix}
\end{proof}
%
Second, we show that at each step of evaluation, the partial input type of $f$
is a refinement of the instantiation of $\vhole{\thole}$.
%
\begin{lem}
\label{lem:resolve-compat}
For all traces
$\trace \defeq \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},\ldots,\triple{e}{\vsu}{\tsu}$,
$\vsub{\ptype{\trace}{f}}{\resolve{\vhole{\thole}}{\vsu}}$.
\end{lem}
\begin{proof}
  By induction on $\trace$.
  %
  In the base case $\trace = \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu}$
  and $\thole$ is trivially a refinement of $\vhole{\thole}$.
  %
  In the inductive case, consider the single-step extension of $\trace$,
  $\trace' = \trace,\triple{e'}{\vsu'}{\tsu'}$.
  %
  We show by case analysis on the evaluation rules that if
  $\vsub{\ptype{\trace}{f}}{\resolve{\vhole{\thole}}{\vsu}}$, then
  $\vsub{\ptype{\trace'}{f}}{\resolve{\vhole{\thole}}{\vsu'}}$.
  %
\end{proof}

@@

\begin{proof}[\textbf{Proof of Theorem~\ref{thm:soundness}}]
%
Suppose $\trace$ witnesses that $f$ gets stuck,
and let $s = \ptype{\trace}{f}$.
We show that \emph{all} types $t$ have stuck-inducing
values by splitting cases on whether $t$ is
compatible with $s$. %the partial type upto $\trace$.
%
\begin{description}
\item [Case \tcompat{s}{t}:]
  Let $\trace = \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},\ldots,\triple{\stuck}{\vsu}{\tsu}$.
  %
  The value $v = \resolve{\vhole{\thole}}{\vsu}$ demonstrates that
  $\eapp{f}{v}$ gets stuck.
\item [Case \tincompat{s}{t}:] By Lemma~\ref{lem:k-stuck}, we can derive
  a $v$ from $\trace$ such that $\hastype{v}{t}$ and $\eapp{f}{v}$ gets
  stuck.
% \item [Case \tincompat{s}{t}:] By Lemma~\ref{lem:k-stuck}, every $v$
%   such that \hastype{v}{t} demonstrates that $\eapp{f}{v}$ gets stuck.
%   % \ES{do we need to say anythign else?}
\end{description}
\end{proof}

>>>>

Next, we formalize the notion of type error witnesses as follows.
%
First, we define a core calculus within which we will work~(\S~\ref{sec:syntax}).
%
Second, we develop a (non-deterministic) operational semantics
for ill-typed programs that precisely defines the notion
of a \emph{witness}~(\S~\ref{sec:semantics}).
%
Third, we formalize and prove a notion of \emph{generality} for
witnesses, which states, intuitively, that if we find a
single witness then for \emph{every possible} type
assignment there exist inputs that are guaranteed to make
the program ``go wrong''~(\S~\ref{sec:soundness}).
%
Finally, we refine the operational semantics into a
\emph{search procedure} that returns concrete (general)
witnesses for ill-typed programs~\S~(\ref{sec:search-algorithm}).
%
We have formalized and tested our semantics and generality theorem
in \textsc{PLT-Redex}~\cite{Felleisen2009-ya}.

@@

\paragraph{Holes}
\label{sec:holes}
%
Recall that a key challenge in our setting is to find witnesses
that are meaningful and do not arise from choosing values from
irrelevant types.
%
We solve this problem by equipping our term language with a
notion of a \emph{hole}, written \vhole{\thole}, which represents
an \emph{unconstrained} value $\ehole$ that may be replaced with
\emph{any} value of an unknown type \thole.
%
Intuitively, the type holes \thole\ can be viewed as type variables
that we will \emph{not} generalize over.
%
A \emph{normalized} value is one that is not a hole,
but which may internally contain holes.
%
For example
$\vnode{\thole}{\vhole{\thole}}{\vleaf{\thole}}{\vleaf{\thole}}$ is a
normalized value.

\paragraph{Substitutions}
%
Our semantics ensure the generality of witnesses by incrementally
\emph{refining} holes, filling in just as much information as is
needed locally to make progress (inspired by the manner in
which SmallCheck uses lazy evaluation~\cite{Runciman2008-ka}).
%
We track how the holes are incrementally filled in, by using
value (resp.\ type) \emph{substitutions} $\vsu$ (resp. $\tsu$)
that map value (resp.\ type) holes to values (resp.\ types).
%
The substitutions let us ensure that we consistently instantiate each
hole with the same (partially defined) value or type, regardless of the
multiple contexts in which the hole appears.
%
This ensures we can report a concrete (and general) witness for any
(dynamically) discovered type errors.

@@

\begin{figure}[ht]
$$
\begin{array}{lcll}
\gensym       & :   & t \times \tsu \rightarrow v \\
\gen{\thole}{\tsu}  & \defeq  & \gen{\subst{\tsu}{\thole}}{\tsu} &  \text{if } \thole \in dom(\tsu) \\
\gen{\tint}{\tsu}   & \defeq  & n &  \text{non-det.} \\
\gen{\tbool}{\tsu}  & \defeq  & b &  \text{non-det.} \\
\gen{\tprod{t_1}{t_2}}{\tsu}  & \defeq  & \epair{\gen{t_1}{\tsu}}{\gen{t_2}{\tsu}} & \\ % \begin{subarray}{l} v_1 = \gen{t_1}{\tsu}\\ v_2 = \gen{t_2}{\tsu}\end{subarray} \\
\gen{\ttree{t}}{\tsu}  & \defeq  & tr &  \text{non-det.} \\
\gen{\tfun}{\tsu}   & \defeq & \efun{x}{\vhole{\thole}} &  \text{\ehole, \thole are fresh} \\
\gen{\thole}{\tsu}  & \defeq & \vhole{\thole} & \text{\ehole is fresh} \\
\end{array}
$$
\caption{Generating values}
\label{fig:gen}
\end{figure}

@@

\paragraph{Recursion}
%
% \RJ{this para appears out of nowhere --- non-sequitur}
Our semantics lacks a built-in $\mathtt{fix}$ construct
for defining recursive functions, which may surprise
the reader.
%
Fixed-point operators often cannot be typed in static type
systems, but our system would simply approximate its type
as $\tfun$, apply it, and move along with evaluation.
%
Thus we can use any of the standard fixed-point operators
and do not need a built-in recursion construct.

@@

\paragraph{Preservation}
%
We prove two preservation lemmas. First, we show that each evaluation
step refines the partial input type of $f$, thus preserving type
compatibility.
%
\begin{lem}
\label{lem:vsu-ext}
If  $\trace \defeq \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},\ldots,\triple{e}{\vsu}{\tsu}$
and $\trace' \defeq \trace, \step{e}{\vsu}{\tsu}{e'}{\vsu'}{\tsu'}$
  (\ie\ $\trace'$ is a single-step extension of $\trace$)
% and $\trace'$ is a single-step extension of $\trace$
and $\ptype{\trace}{f} \neq \ptype{\trace'}{f}$
% and $\resolve{\thole}{\tsu} \neq \resolve{\thole}{\tsu'}$
then $\tsu' = \tsu[\thole_1 \mapsto t_1] \ldots [\thole_n \mapsto t_n]$.
\end{lem}
\begin{proof}
  By case analysis on the evaluation rules.
  %
  $\thole$ does not change, so if the partial input types differ then
  $\tsu \neq \tsu'$.
  %
  Note that only \forcesym\ can change \tsu,
  %, and it can only do so via \unifysym
  via \unifysym, which can only extend \tsu.
  % and only the first case,
  % when we extend \vsu\ with $[\ehole \mapsto v]$,
  % does.
  %
  % \ES{fill in cases in the appendix}
\end{proof}
%
Second, we show that at each step of evaluation, the partial input type of $f$
is a refinement of the instantiation of $\vhole{\thole}$.
%
\begin{lem}
\label{lem:resolve-compat}
For all traces
$\trace \defeq \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},\ldots,\triple{e}{\vsu}{\tsu}$,
$\vsub{\ptype{\trace}{f}}{\resolve{\vhole{\thole}}{\vsu}}$.
\end{lem}
\begin{proof}
  By induction on $\trace$.
  %
  In the base case $\trace = \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu}$
  and $\thole$ is trivially a refinement of $\vhole{\thole}$.
  %
  In the inductive case, consider the single-step extension of $\trace$,
  $\trace' = \trace,\triple{e'}{\vsu'}{\tsu'}$.
  %
  We show by case analysis on the evaluation rules that if
  $\vsub{\ptype{\trace}{f}}{\resolve{\vhole{\thole}}{\vsu}}$, then
  $\vsub{\ptype{\trace'}{f}}{\resolve{\vhole{\thole}}{\vsu'}}$.
  %
\end{proof}

@@

\begin{proof}[\textbf{Proof of Theorem~\ref{thm:soundness}}]
%
Suppose $\trace$ witnesses that $f$ gets stuck,
and let $s = \ptype{\trace}{f}$.
We show that \emph{all} types $t$ have stuck-inducing
values by splitting cases on whether $t$ is
compatible with $s$. %the partial type upto $\trace$.
%
\begin{description}
\item [Case \tcompat{s}{t}:]
  Let $\trace = \triple{\eapp{f}{\vhole{\thole}}}{\emptysu}{\emptysu},\ldots,\triple{\stuck}{\vsu}{\tsu}$.
  %
  The value $v = \resolve{\vhole{\thole}}{\vsu}$ demonstrates that
  $\eapp{f}{v}$ gets stuck.
\item [Case \tincompat{s}{t}:] By Lemma~\ref{lem:k-stuck}, we can derive
  a $v$ from $\trace$ such that $\hastype{v}{t}$ and $\eapp{f}{v}$ gets
  stuck.
% \item [Case \tincompat{s}{t}:] By Lemma~\ref{lem:k-stuck}, every $v$
%   such that \hastype{v}{t} demonstrates that $\eapp{f}{v}$ gets stuck.
%   % \ES{do we need to say anythign else?}
\qedhere
\end{description}
\end{proof}
